Predict Disease Outcome by Community Structures

### 2016-08-16 code/HumanCdGF_microbiota_predict.Rmd
Add RF prediction of mobundity

```{r, message=FALSE}
pack_used <- c('randomForest','miscTools','ggplot2','reshape2', 'pROC')
for (dep in pack_used){
  if (dep %in% installed.packages()[,"Package"] == FALSE){
    install.packages(as.character(dep), repos = 'http://cran.us.r-project.org', 
                     quiet=TRUE);
  }
  library(dep, verbose=FALSE, character.only=TRUE)
}
```

```{r}
setwd("~/Documents/Github/Schubert_humanCdGF_XXXX_2016")

# read in files
meta_file   <- read.table('data/process/human_CdGF_metadata.txt', sep = '\t',
                        header = T, row.names = 'group')
shared_file <- read.table('data/process/human_CdGF.an.unique_list.0.03.subsample.shared',
                          sep='\t',header = T, row.names='Group')
tax_file <- read.table('data/mothur/gf_cdiff.trim.contigs.good.unique.good.filter.unique.precluster.pick.pick.pick.an.unique_list.0.03.cons.taxonomy', 
                       sep='\t',header = T, row.names = 'OTU')
#source('code/tax_level.R')

# add log transformed CFU
meta_file$log_cfu <- log10(meta_file$cdiff_cfu + 1)
# remove cecum samples
meta_file <- meta_file[! meta_file$sample_type == 'cecum', ]

@@

ggplot(n_feat_day_euth, aes(x=Features..n., y=RSQ, )) + geom_line() +
  labs(title= 'R^2 from RF with top n features\n
               Day euthanized C Diff CFU from Day 0 Community', 
       x = 'Features (n)', y = 'R^2') + theme_bw()

@@

##Can we predict moribundity with the community structure at Day 0?
```{r}
##Fix to be categorical, and not regression
Predict_early_euth_df <- data.frame(Euth_Early = ifelse(meta_file[meta_file$day %in% 0,
                                                                  'Euthanized'] %in% 10,
                                                        'No','Yes'), 
                                    row.names = meta_file[meta_file$day %in% 0, 'Mouse_ID'])
Predict_early_euth_df <- merge(Predict_early_euth_df, rel_abund_stool, by='row.names')
Predict_early_euth_df <- Predict_early_euth_df[,!(colnames(Predict_early_euth_df) %in% "Row.names")]
rf_early_euth <- randomForest(as.factor(Euth_Early) ~ ., data = Predict_early_euth_df, 
                              importance = TRUE, ntree = n_trees)
# write(paste("binary", ncol(rel_abund_stool), rf_early_euth$err.rate[n_trees], "\t"), file="data/process/random_forest.data", append=T)
#varImpPlot(rf_early_euth)
#prediction_euth <- as.numeric(predict(rf_early_euth, Predict_early_euth_df[,-1]))
#ROC_euth <- roc(as.numeric(Predict_early_euth_df$Euth_Early), prediction_euth)
#plot(ROC_euth)
```

>>>>

```{r, message=FALSE}
pack_used <- c('randomForest','miscTools','ggplot2','reshape2', 'pROC', 'knitr')
for (dep in pack_used){
  if (dep %in% installed.packages()[,"Package"] == FALSE){
    install.packages(as.character(dep), repos = 'http://cran.us.r-project.org', 
                     quiet=TRUE);
  }
  library(dep, verbose=FALSE, character.only=TRUE)
}
```

```{r}
setwd("~/Documents/Github/Schubert_humanCdGF_XXXX_2016/code")

# read in files
meta_file   <- read.table('../data/process/human_CdGF_metadata.txt', sep = '\t',
                        header = T, row.names = 'group')
shared_file <- read.table('../data/process/human_CdGF.an.unique_list.0.03.subsample.shared',
                          sep='\t',header = T, row.names='Group')
tax_file <- read.table('../data/mothur/gf_cdiff.trim.contigs.good.unique.good.filter.unique.precluster.pick.pick.pick.an.unique_list.0.03.cons.taxonomy',
                       sep='\t',header = T, row.names = 'OTU')
#source('tax_level.R')

# add log transformed CFU
meta_file$log_cfu <- log10(meta_file$cdiff_cfu + 1)
# remove cecum samples
meta_file <- meta_file[! meta_file$sample_type == 'cecum', ]
# remove unused cage P2 and non 431 cdiff
meta_file <- meta_file[! meta_file$cage_id %in% c('P2', 'CON', 'CON2', 'CON3'), ]

@@

ggplot(n_feat_day_euth, aes(x=Features..n., y=RSQ)) + geom_line() +
  labs(title= 'R^2 from RF with top n features\n
               Day euthanized C Diff CFU from Day 0 Community', 
       x = 'Features (n)', y = 'R^2') + theme_bw()

@@

##Can we predict moribundity with the community structure at Day 0?
```{r}
Predict_early_euth_df <- data.frame(Euth_Early = ifelse(meta_file[meta_file$day %in% 0,
                                                                  'Euthanized'] %in% 10,
                                                        'No','Yes'), 
                                    row.names = meta_file[meta_file$day %in% 0, 'Mouse_ID'])
Predict_early_euth_df <- merge(Predict_early_euth_df, rel_abund_stool, by='row.names')
mouse_list <- Predict_early_euth_df$Row.names
Predict_early_euth_df <- Predict_early_euth_df[,!(colnames(Predict_early_euth_df) %in% "Row.names")]
rf_early_euth <- randomForest(as.factor(Euth_Early) ~ ., data = Predict_early_euth_df, 
                              importance = TRUE, ntree = n_trees)

unpredicted <- mouse_list[Predict_early_euth_df$Euth_Early != rf_early_euth$predicted]
rf_euth_import <- data.frame(importance = head(
  sort(rf_early_euth$importance[ , 'MeanDecreaseAccuracy'], decreasing = T), 30))
ROC_euth <- roc(as.numeric(Predict_early_euth_df$Euth_Early), as.numeric(rf_early_euth$predicted))
plot(ROC_euth)

ggplot(data = rf_euth_import, aes(x = factor(rownames(rf_euth_import)), y = importance)) + 
        geom_point() + scale_x_discrete(limits = rev(rownames(rf_euth_import))) +
        labs(title = 'OTU Importance in Prediction of Day Euthanized from Day 0 Community',
        x= '', y = '% Increase in MSE') + theme_bw() + coord_flip()
```

Confusion Matrix  
`r kable(rf_early_euth$confusion)`

OOB error rate = `r 100*tail(rf_early_euth$err.rate[ , 1], 1)`

Mice which were euthanized early, but the model predicts the will were not (Mouse Tag - Cage):
`r unpredicted`
